---
title: "FitBit Data Analysis"
author: "Alan Johnson"
date: "March 2016"
output: html_document
---
##About
This was a project to analyze the FitBit data given for Project 1 in Reproducable Research.


##Overview
The purpose of this was to get the following data:
1. Code for reading in the dataset and/or processing the data
2. Histogram of the total number of steps taken each day
3. Mean and median number of steps taken each day
4. Time series plot of the average number of steps taken
5. The 5-minute interval that, on average, contains the maximum number of steps
6. Code to describe and show a strategy for imputing missing data
7. Histogram of the total number of steps taken each day after missing values are imputed
8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
9. All of the R code needed to reproduce the results (numbers, plots, etc.) in the report

##Load and process the original data

Load data

```{r unzip, cache=TRUE}
unzip("activity.zip")
```

```{r reading, cache=TRUE}
data <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
Sum by day, create histogram, and calculate mean and median.
```{r} 
stepsbyday <- aggregate(steps ~ date, data, sum)
hist(stepsbyday$steps, main = paste("Total Steps Each Day"), col="orange", xlab="Number of Steps")
rmean <- mean(stepsbyday$steps)
rmedian <- median(stepsbyday$steps)
```

## What is the average daily activity pattern?

```{r}
stepsbyinterval <- aggregate(steps ~ interval, data, mean)

plot(stepsbyinterval$interval,stepsbyinterval$steps, type="l", xlab="Interval", ylab="Number of Steps",main="Average Number of Steps per Day by Interval")

maxinterval <- stepsbyinterval[which.max(stepsbyinterval$steps),1]
```

## Impute missing values. Compare imputed to non-imputed data.
Missing data needed to be imputed. 
Missing values were imputed by inserting the average for each interval. 
```{r}
incomplete <- sum(!complete.cases(data))
imputeddata <- transform(data, steps = ifelse(is.na(data$steps), stepsbyinterval$steps[match(data$interval, stepsbyinterval$interval)], data$steps))
```

Zeroes were imputed for 10-01-2012 because it was the first day 
```{r}
imputeddata[as.character(imputeddata$date) == "2012-10-01", 1] <- 0
```

Recount total steps by day and create Histogram. 
```{r}
stepsbyday1 <- aggregate(steps ~ date, imputeddata, sum)
hist(stepsbyday1$steps, main = paste("Total Steps Each Day"), col="orange", xlab="Number of Steps")

#Create Histogram to show difference. 
hist(stepsbyday$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps", add=T)
legend("topright", c("Imputed", "Non-imputed"), col=c("orange", "orange"), lwd=10)
```

Calculate new mean and median for imputed data. 
```{r}
rmean1 <- mean(stepsbyday1$steps)
rmedian1 <- median(stepsbyday1$steps)
```

Calculate difference between imputed and non-imputed data.
```{r}
meandiff <- rmean1 - rmean
meddiff <- rmedian1 - rmedian
```

Calculate total difference.
```{r}
totaldiff <- sum(stepsbyday1$steps) - sum(stepsbyday$steps)
```

## Are there differences in activity patterns between weekdays and weekends?
Created a plot to compare and contrast number of steps between the week and weekend. 
``` {r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")
imputeddata$dow = as.factor(ifelse(is.element(weekdays(as.Date(imputeddata$date)),weekdays), "Weekday", "Weekend"))

stepsbyinterval1 <- aggregate(steps ~ interval + dow, imputeddata, mean)

library(lattice)

xyplot(stepsbyinterval1$steps ~ stepsbyinterval1$interval|stepsbyinterval1$dow, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")

```
